The \dealii{} Library, Version 9.0

### 2018-04-04 9.0/paper.tex
list of changes url

\dealii{} version 9.0.0 was released April X, 2018. This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.0. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

@@

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{in
the file that lists all changes for this release}, see \cite{changes90}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

@@

\marginpar{OLD}
The 9.0 release includes around XYZ
\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{
incompatible changes}; see \cite{changes90}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions, and the majority change internal
interfaces that are not usually used in external applications. However, three
incompatible changes are worth mentioning:
\begin{itemize}
  \item High-order Lagrange elements, both continuous \verb!FE_Q! and
    discontinuous \verb!FE_DGQ! types, now use the nodal points of the
    Gauss-Lobatto quadrature formula as support points by default, rather than the
    previous equidistant ones. For cubic polynomials and higher, the point
    distribution has thus changed and, consequently, the entries in
    solution vectors will be different compared to previous
    versions of \dealii{}. Note, however, that using the Gauss-Lobatto points as nodal
    points results in a much more stable interpolation, including better
    iteration counts in most iterative solvers.
  \item
    The library no longer instantiates template classes with \texttt{long
    double}. These were rarely used, but took up a significant
    fraction of compile and link time, as well as library
    size. Application programs can, however, still instantiate all
    template classes with \texttt{long
    double} as long as they include the corresponding \texttt{.templates.h}
    header files.
  \item
    The \texttt{ParameterGUI} has been moved to a separate repository.
\end{itemize}

>>>>

\dealii{} version 9.0.0 was released April X, 2018. This paper provides an
overview of the new features of this major release and serves as a citable
reference for the \dealii{} software library version 9.0. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

@@

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
in the file that lists all changes for this release}, see \cite{changes90}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

@@

\marginpar{OLD}
The 9.0 release includes around XYZ
\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
incompatible changes}; see \cite{changes90}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions, and the majority change internal
interfaces that are not usually used in external applications. However, three
incompatible changes are worth mentioning:
\begin{itemize}
  \item High-order Lagrange elements, both continuous \verb!FE_Q! and
    discontinuous \verb!FE_DGQ! types, now use the nodal points of the
    Gauss-Lobatto quadrature formula as support points by default, rather than the
    previous equidistant ones. For cubic polynomials and higher, the point
    distribution has thus changed and, consequently, the entries in
    solution vectors will be different compared to previous
    versions of \dealii{}. Note, however, that using the Gauss-Lobatto points as nodal
    points results in a much more stable interpolation, including better
    iteration counts in most iterative solvers.
  \item
    The library no longer instantiates template classes with \texttt{long
    double}. These were rarely used, but took up a significant
    fraction of compile and link time, as well as library
    size. Application programs can, however, still instantiate all
    template classes with \texttt{long
    double} as long as they include the corresponding \texttt{.templates.h}
    header files.
  \item
    The \texttt{ParameterGUI} has been moved to a separate repository.
\end{itemize}
